---
title: "Kaggle Part 2"
author: "Angel Claudio"
date: "3/19/2020"
output: html_document
---

# Kaggle DS survery 2019 - PART TWO

## R LIBRARY SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(odbc)
library(DBI)
library(ggmosaic)
```

## Open Connection to Database

```{r db-authenticate, eval=T, echo=T, message=FALSE, Warning=F}
conn <- dbConnect(odbc(),
                  Driver = "SQL Server",
                  Server = "spsangelclaudio.database.windows.net",
                  Database = "DataScienceValue",
                  UID = "spsadmin",
                  PWD = "fakepw")
```

## Get View Object for Use

```{r}
#kaggle_data <- tbl(conn, "uvwParticipantData")
kaggle_data <- dbGetQuery(conn, 'Select * from uvwParticipantData')
```

```{r}

```

## Check for outliers

There are no possible numeric outliers as users did not input numbers by hand but instead chose a bin provided by the survey writers.

## Cleaning Salary data


I will clean salary data to make it numeric, allowing it to be easily plotted.

```{r}

##This function takes numeric bins and returns the mean of the bin, allowing us to more easily order/graph them.
clean_numeric_bins <- function(series){
  series <- str_remove_all(series, '[,><+/s]') %>%  str_split('-')
  series <- sapply(series, as.integer) %>% sapply(mean) %>% sapply(round)
  series
}



df_basic <- kaggle_data
df_basic$Salary <- clean_numeric_bins(df_basic$Salary)
```

## Do age and gender affect compensation?

### Univariate Plots

First we examine the distributions of age, gender, and compensation on their own.

```{r}
df_basic %>% ggplot(aes(x = Age, fill = Gender)) + geom_histogram(alpha = 0.5, stat = 'count', position = 'identity')

mean(df_basic$Gender == 'Male')
```

The first thing we notice is that the sample is extremely skewed, made up of almost predominantly males. In fact, almost 84% of those surveyed are male. The most common age group is 25-29, and it appears the median age is likely somewhere around 30.

```{r}
df_basic %>% ggplot(aes(factor(Salary))) + geom_bar(stat = 'count') + theme(axis.text.x = element_text(angle = 90))
```

We see that the mode salary is extremely low, 500$. This is unexpected, and likely related to people filling out the survey who are not yet working as data scientists still labeling themself as a data scientist.

### Bivariate Plots

Now we will examine how age and compensation and gender and compensation vary together.

I plot these using mosaic plots, where the area of a rectangle represents the proportion of people falling into both groups.

```{r}
df_basic %>%
  ggplot() + 
  geom_mosaic(aes(x = product(Age, Salary), fill = Age), na.rm = TRUE) + xlab('Salary') + ylab('Age') + ggtitle('Salary by Age Group') + theme(axis.text = element_text(size = 6), axis.text.x = element_text(angle = 90))

df_basic %>%
  ggplot() + 
  geom_mosaic(aes(x = product(Gender, Salary), fill = Gender), na.rm = TRUE) + xlab('Salary') + ylab('Gender') + ggtitle('Salary by Gender') + theme(axis.text = element_text(size = 6), axis.text.x = element_text(angle = 90))

df_basic %>% ggplot(aes(x = Gender, y = Salary, fill = Gender)) + geom_boxplot()

df_basic %>% ggplot(aes(x = Age, y = Salary, fill = Age)) + geom_boxplot()
```

There is a clear distinction between male and female salaries, and a clear upward trend between age and salary.


## Education, Country, and Experience vs. Salary

### Univariate Plots

First we get a feeling for each variable on its own via a bar plot. I am going to apply the same approach to years of experience to make it more useable in a plot.

```{r, warning = FALSE}
df_basic$Education <- df_basic$Education %>% str_replace_all('Some collegeuniversity study without earning a bachelors', 'Some college') %>% str_replace_all('No formal education past high school', 'High school')

df_basic %>% ggplot(aes(x = Education)) + geom_bar(stat = 'count') + theme(axis.text.x = element_text(angle = 90))

##TODO REORDER BY COUNT, fix off tick labels
df_basic %>% ggplot(aes(x = Country)) + geom_bar(stat = 'count') + theme(axis.text.x = element_text(angle = 90))

df_basic$codeExperience<- clean_numeric_bins(df_basic$YearsOfCodeExperience)
df_basic %>% ggplot(aes(x = factor(codeExperience))) + geom_bar(stat = 'count') + theme(axis.text.x = element_text(angle = 90))
```

We see that the most common level of education is a master's degree, the most common country is the United States, although India is extremely close, and the most common level of coding experience is approximately 4 years.

### Bivariate Plots

```{r, warning = FALSE}
df_basic %>%
  ggplot() + 
  geom_mosaic(aes(x = product(Education, Salary), fill = Education), na.rm = TRUE) + xlab('Salary') + ylab('Education') + ggtitle('Salary by Education') + theme(axis.text = element_text(size = 6), axis.text.x = element_text(angle = 90))

india_us <- df_basic %>% filter(Country == 'United States of America'|Country == 'India')

india_us %>%
  ggplot() + 
  geom_mosaic(aes(x = product(Country, Salary), fill = Country), na.rm = TRUE) + xlab('Salary') + ylab('Country') + ggtitle('Salary by Country') + theme(axis.text = element_text(size = 6), axis.text.x = element_text(angle = 90))

df_basic %>%
  ggplot() + 
  geom_mosaic(aes(x = product(factor(codeExperience), Salary), fill = factor(codeExperience)), na.rm = TRUE) + xlab('Salary') + ylab('Coding Experience') + ggtitle('Salary by Coding Experience') + theme(axis.text = element_text(size = 6), axis.text.x = element_text(angle = 90))
```

We find mostly what we would expect. Compensation goes up with education, is HIGHLY affected by country of origin, as US compensation greatly outpaces India's, and goes up with years of coding experience.

```{r, warning = FALSE}
df_basic %>% ggplot(aes(Education, Salary, fill = Education)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90), legend.position = 'None')

india_us %>% ggplot(aes(Country, Salary, fill = Country)) + geom_boxplot() + theme(legend.position = 'None')

df_basic %>% ggplot(aes(factor(codeExperience), Salary, fill = factor(codeExperience))) + geom_boxplot() + xlab('Years of Coding Experience') + theme(legend.position = 'None')
```

## Beyond the basic dataframe

We have lots of information broken down into many sub dataframes. I will attempt to access this information via a function rather than doing each dataframe individually.

```{r}
rm(india_us)

df_list <- list('df_ide', 'df_lang', 'df_viz', 'df_ml_algo', 'df_ml_tool', 'df_cv', 'df_nlp', 'df_cloud', 'df_cloud_prod', 'df_big_data', 'df_cloud_ml', 'df_db')
```

```{r, warning = FALSE}
### Function to summarize the smaller dataframes into medians and counts
summarize_df <- function(df_name){
  query <- paste('Select * from ', df_name, collapse = '')
  func_df <- dbGetQuery(conn, query)
  names(func_df) <- names(func_df) %>% str_remove_all("[/,-]") %>% str_replace_all('\\+', 'plus')
  titles <- c()
  med <- c()
  counts <- c()
  mega_df <- cbind(df_basic, func_df)
  
  colNames <- names(func_df)
  
  ## iterate through the columns of the dataframe
  ## each column is one skill
  for (j in colNames){
    ## Calculate the median and count for each column
    #mega_df[j] <- mega_df[j] %>% str_remove_all("[\\+,/,-]")
    meds <- mega_df %>% group_by_(j) %>% summarize(med = median(Salary, na.rm = T), count = n())
    titles <- append(titles, meds[[1]])
    med <- append(med, meds[[2]])
    counts <- append(counts, meds[[3]])
  }
  ##Create a dataframe of all skill names, medians, counts
  return(data.frame('skill' = titles, 'medians' = med, 'counts' = counts))
}
```

```{r, warning = FALSE}

##Plot side dataframes that are in slightly different format

df_prime_tool <- dbGetQuery(conn, 'Select * from df_prime_tool')
cbind(df_basic, df_prime_tool) %>% ggplot(aes(x = str_wrap(primary_analysis_tool, 12), y = Salary, fill = primary_analysis_tool)) + geom_boxplot() + theme(axis.text.x = element_text(size = 6), legend.text = element_text(size = 6)) + xlab('Primary Analysis Tool') + theme(legend.position = 'none')

df_ml_xp <- dbGetQuery(conn, 'Select * from df_ml_xp')
df_ml_xp$ml_exp_years <- clean_numeric_bins(df_ml_xp$ml_exp_years)
cbind(df_basic, df_ml_xp) %>% ggplot(aes(x = factor(ml_exp_years), y = Salary, fill = factor(ml_exp_years))) + geom_boxplot() + xlab('Years of ML Experience') + theme(legend.position = 'None')

##Plot other side dataframes using function created.
for (i in df_list){
  winnerdf <- summarize_df(i)
  winnerdf <- winnerdf %>% filter(!(skill %in% c('None', 'Other'))) %>% filter(!(is.na(skill)))
  
  plots <- winnerdf %>% ggplot(aes(counts, medians, color = str_wrap(skill, 15))) + geom_point() + theme(legend.text = element_text(size = 5), legend.title = element_blank()) + xlab('Popularity (counts)') +ylab('Median Salary')
  print(plots)
}

```

### Breakdown of Results

