---
title: "LQ- Rough Draft Scraped Data And Visuals"
author: "Layla Quinones"
date: "3/14/2020"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, hide = TRUE}
library(rvest)
library(stringi)
library(tidyverse)
library(RCurl)
library(RColorBrewer)
library(xml2)
library(kableExtra)
library(Stack)
```
# Data Analysis

## Core Data Science Skills Professionals Have and Want

A group of 1500 Data Scientists were asked the following questions:

1. Which skills / knowledge areas do you currently have (at the level you can use in work or research)?

2. Which skills do you want to add or improve?
```{r, warning = FALSE}
#Clean KD Nuggets DS Table 
urlTen <- "https://www.kdnuggets.com/2019/09/core-hot-data-science-skills.html"

#Parse
urlTen <- read_html(urlTen)

#Reads the tables in the html page
KdTables <- html_nodes(urlTen, "table") %>%
  html_table(fill = TRUE)

#Tables
kdTableOne <- KdTables[[1]] %>% mutate(Type = as.factor("Established Skills"))
kdTableTwo <- KdTables[[2]] %>% mutate(Type = as.factor("New Skills"))
kdTableThree <- KdTables[[3]] %>% mutate(Type = as.factor("Other"))

#Stack the Data Frames
combinedKdTable <- Stack(kdTableOne, kdTableTwo)
combinedKdTable <- Stack(combinedKdTable, kdTableThree)

#Rename Columns
combinedKdTable <- rename(combinedKdTable, "want" = "%Want", "have" = "%Have","wantHaveRatio" ="%Want/%Have") %>% gather("Category", "Percent", "have":"wantHaveRatio")

#Take away % and cast to numeric
combinedKdTable$Percent <- as.numeric(unlist(str_remove_all(combinedKdTable$Percent, "\\%")))
#Put back into wide format for plotting
combinedKdTable<- spread(combinedKdTable, Category, Percent) %>% select(Skill, want,have, Type)

#plotting have vs want
ggplot(combinedKdTable, aes(x=have, y=want, color = Type)) +
  scale_color_brewer(palette="Dark2") +
  labs(y = "Percent Want", x ="Percent Have", title = "Skills Data Scientists Want vs Have (Clustering)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_label(label=combinedKdTable$Skill, nudge_x = 0.25, nudge_y = 0.25, check_overlap = T)
  
```
> Here we can see clearly that the skills identified in this survey falls into three categories: skills that are well established that have a high percentage of respondants that have them and a low percentage of respondants who want them(green); skills that are in high demand with a high percentage of respondents reporting they want that skill and do not have them(orange); and skills that very few people want and very few people have (purple). Below each category is analyzed in detail.

### Well Establish Data Science Skills 

```{r}
#Clean Up the table for analysis
kdTableOne <- rename(kdTableOne, "want" = "%Want", "have" = "%Have","wantHaveRatio" ="%Want/%Have") %>% gather("Category", "Percent", "have":"wantHaveRatio")

kdTableOne$Percent <- as.numeric(unlist(str_remove_all(kdTableOne$Percent, "\\%")))

wantHave <- spread(kdTableOne, Category, Percent) %>% select(Skill, want,have)
wantHave <- gather(wantHave,"Category", "Percent", "have":"want")

#Table comparing Want and Have
kable(wantHave) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Visualization of Want and Have
ggplot(wantHave, aes(fill=Category, y=Percent, x=reorder(Skill,Percent))) + 
  scale_fill_brewer(palette="Set3") +
  geom_col(position="dodge") +
  labs(y = "Percent of Respondants", x ="Established Data Science Skills", title = "Established Skills Data Scientists Want or Have") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 skills that most data scientists reported having are Python, Data Visualization and Excel.

> The top 3 skills that data scientists reported wanting is Machine Learning, Python and Statistics.

```{r}
#Table with Ratio of want/have
propWantHave <- spread(kdTableOne, Category, Percent) %>% select(Skill, wantHaveRatio)

#Visualization
ggplot(propWantHave, aes(y=wantHaveRatio, x=reorder(Skill,wantHaveRatio))) +
  geom_col(fill = "lightgreen", position="dodge") +
  labs(y = "%Want/%Have Ratio", x ="Established DS Skills", title = "Ratio of %Want/%Have DS Skills") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 established skills with the greatest proportion of percent want to percent have is Machine Learning, Scikit-learn and Python, whereas the 3 skills with the lowest proportion was Excel, Critical Thinking and Communication skills (which may indicate skills that data scientists are highly proficent at).

### Emerging Data Science Skills

```{r}
#Clean Up the table for analysis
kdTableTwo <- rename(kdTableTwo, "want" = "%Want", "have" = "%Have","wantHaveRatio" ="%Want/%Have") %>% gather("Category", "Percent", "have":"wantHaveRatio")

kdTableTwo$Percent <- as.numeric(unlist(str_remove_all(kdTableTwo$Percent, "\\%")))

wantHaveTwo <- spread(kdTableTwo, Category, Percent) %>% select(Skill, want,have)
wantHaveTwo <- gather(wantHaveTwo,"Category", "Percent", "have":"want")

#Table comparing Want and Have
kable(wantHaveTwo) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Visualization of Want and Have
ggplot(wantHaveTwo, aes(fill=Category, y=Percent, x=reorder(Skill,Percent))) + 
  scale_fill_brewer(palette="Set2") +
  geom_col(position="dodge") +
  labs(y = "Percent of Respondants", x ="Data Science Skills", title = "Emerging Skills Data Scientists Want or Have") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 "up and coming" skills that most data scientists reported having are Deep Learning, NLP - Text Processing and Unsturctured Data.

> The top 3 "up and coming" skills that data scientists reported wanting is Deep Learning, Tensor Flow and Apache Spark.

```{r}
#Table with Ratio of want/have
propWantHaveNew <- spread(kdTableTwo, Category, Percent) %>% select(Skill, wantHaveRatio)

#Visualization
ggplot(propWantHaveNew, aes(y=wantHaveRatio, x=reorder(Skill,wantHaveRatio))) +
  geom_col(fill = "lightpink", position="dodge") +
  labs(y = "%Want/%Have Ratio", x ="New DS Skills", title = "Ratio of %Want/%Have for Emerging DS Skills") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 emerging skills with the greatest proportion of percent want to percent have is Pytorch, Scala and "Python"Other Big Data Tools", whereas the 3 skills with the lowest proportion was Unstructured Data, Kaggle and NLP Text Processing.

### Other DS Skills
```{r}
#Clean Up the table for analysis
kdTableThree <- rename(kdTableThree, "want" = "%Want", "have" = "%Have","wantHaveRatio" ="%Want/%Have") %>% gather("Category", "Percent", "have":"wantHaveRatio")

kdTableThree$Percent <- as.numeric(unlist(str_remove_all(kdTableThree$Percent, "\\%")))

wantHaveThree <- spread(kdTableThree, Category, Percent) %>% select(Skill, want,have)
wantHaveThree <- gather(wantHaveThree,"Category", "Percent", "have":"want")

#Table comparing Want and Have
kable(wantHaveThree) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Visualization of Want and Have
ggplot(wantHaveThree, aes(fill=Category, y=Percent, x=reorder(Skill,Percent))) + 
  scale_fill_brewer(palette="Accent") +
  geom_col(position="dodge") +
  labs(y = "Percent of Respondants", x =" 'Other' Data Science Skills", title = "'Other' Skills Data Scientists Want or Have") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 "other" skills that most data scientists reported having are Software Engineering, XGBoost and Java.

> The top 3 "other" skills that data scientists reported wanting are the same 3 (Software Engineering, XGBoost and Java).


```{r}
propWantHaveOther <- spread(kdTableThree, Category, Percent) %>% select(Skill, wantHaveRatio)

#Visualization
ggplot(propWantHaveOther, aes(y=wantHaveRatio, x=reorder(Skill,wantHaveRatio))) +
  geom_col(fill = "lightblue", position="dodge") +
  labs(y = "%Want/%Have Ratio", x ="Other DS Skills", title = "Ratio of %Want/%Have for 'Other' DS Skills") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> The top 3 emerging skills with the greatest proportion of percent want to percent have is Pytorch, Scala and "Python"Other Big Data Tools", whereas the 3 skills with the lowest proportion was Unstructured Data, Kaggle and NLP Text Processing.

### Employment

```{r}
#Employment Types
kdEmpType <- html_nodes(urlTen, ".three_ul li") %>%
  html_text() %>%
  .[1:11]

#Get Columns for DF
colData <- unlist(str_extract_all(kdEmpType, '.*,'))
colDataLeft <- unlist(str_remove_all(colData, ','))
colDataRight <- unlist(str_extract_all(kdEmpType, '\\d+.\\d+'))

#CReate DFs
#Employment Type
EmpType <- data.frame("EmpType" = as.factor(colDataLeft[1:5]), "Percent" = as.numeric(colDataRight[1:5]))

#Regional Distribution
regDis<- data.frame("Region" = as.factor(colDataLeft[6:11]), "Percent" = as.numeric(colDataRight[6:11]))

#display Tables
kable(EmpType) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Visualization
ggplot(EmpType, aes(y=Percent, x=reorder(EmpType, Percent))) +
  geom_col(fill = "red", position="dodge") +
  labs(y = "Percent", x ="Employment Type", title = "Percent of DS in Each Emlpoyment Type") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> A vast majority of Data Scientists are Self-Employed or in Industry.

```{r}
#Region and Percent Jobs
kable(regDis) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Visualization
ggplot(regDis, aes(y=Percent, x=reorder(Region, Percent))) +
  geom_col(fill = "pink", position="dodge") +
  labs(y = "Percent", x ="Region", title = "Percent of DS Jobs In Each Region") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

>Top 3 regions hiring Data Scientists are US/ Canada, Europe and Asia. Not surprising.

## KD Poll Software

The 20th annual KDnuggets Software Poll had over 1,800 participants. Data is in percent of voters.

`topSF`: Top Analytics/DS/ML Software in 2019 - 2017

`topUse`: Major Analytics/Data Science/ML Software with the largest increase in usage

`declineUse`: Major Analytics/Data Science Platform with the largest decline in usage

`dlPlat`:Major Deep Learning Platforms

`bdTools`: Big Data Tools

`progLang`: Programming Languages

### Top Software

```{r}
#Clean KD Nuggets DS Table 
urlEleven <- "https://www.kdnuggets.com/2019/05/poll-top-data-science-machine-learning-platforms.html"

#Parse
urlEleven <- read_html(urlEleven)

#Reads the tables in the html page
kdTableFour <- html_nodes(urlEleven, "table") %>%
  html_table(fill = TRUE)

topSF <- kdTableFour[[1]]

kable(topSF) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
topSF <- rename(topSF, "2019" = "2019% share", "2018" = "2018% share","2017" ="2017% share") %>% gather("Year", "Percent", "2019":"2017")

topSF$Percent <- as.numeric(unlist(str_remove_all(topSF$Percent, "\\%")))
topSF$Year <- as.numeric(topSF$Year)

ggplot(topSF) +
  geom_line(aes(x=Year, y=Percent, col = Software)) +
  labs(title= "Top Data Science Tools", x = "Year", y = "Percent of People Using Tool") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

> Python is highest all three years.

> R is in a decline

> RapidMiner is the highest increase in software.

> Tensorflow, Anaconda and Keras are increaing in use.

### Largest Increase

```{r}
topUse <- kdTableFour[[2]]

kable(topUse) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
topUse <- rename(topUse, "2019" = "2019% share", "2018" = "2018% share","change" ="% change") %>% gather("Year", "Percent", "2019":"change")

topUse$Percent <- as.numeric(unlist(str_remove_all(topUse$Percent, "\\%")))

topUse <- filter(topUse, Year == "change")

#Visualization of Want and Have
ggplot(topUse, aes(y=Percent, x=reorder(Software,Percent))) +
  geom_col(fill = "blue") +
  labs(y = "Percent Increase", x ="Data Science Tools", title = "Software With Largest Increase in Use (2018 - 2019)") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> This Data can indicate software tools/skills that will be in demand in future years.

### Top Software Deline Use

```{r}
declineUse <- kdTableFour[[3]]

kable(declineUse) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
declineUse <- rename(declineUse, "2019" = "2019% share", "2018" = "2018% share","change" ="% change") %>% gather("Year", "Percent", "2019":"change")

declineUse$Percent <- as.numeric(unlist(str_remove_all(declineUse$Percent, "\\%")))

declineUse <- filter(declineUse, Year == "change")

#Visualization of Want and Have
ggplot(declineUse, aes(y=Percent, x=reorder(Platform,Percent))) +
  geom_col(fill = "blue") +
  labs(y = "Percent Decline", x ="Data Science Tools", title = "Software With Largest Decrease in Use (2018 - 2019)") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> Indicates tools/skills that will be obselete in the future.

### Deep Learning Platform Trends

```{r}
dlPlat <- kdTableFour[[4]]

kable(dlPlat) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
dlPlat <- rename(dlPlat, "2019" = "2019% share", "2018" = "2018% share","change" ="% change") %>% gather("Year", "Percent", "2019":"change")

dlPlat$Percent <- as.numeric(unlist(str_remove_all(dlPlat$Percent, "\\%")))

dlPlat <- filter(dlPlat, dlPlat$Year == "change")

#Visualization of Want and Have
ggplot(dlPlat, aes(y=Percent, x=reorder(Platform,Percent))) +
  geom_col(fill = "lightgreen") +
  labs(y = "Percent Change", x ="Data Science Tools", title = "Deep Learning Platforms Trends (2018 - 2019)") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```

> Platforms that may or may not be in demand

```{r, warning == FALSE}
bdTools <- kdTableFour[[5]]

kable(bdTools) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
bdTools <- rename(bdTools, "2019" = "2019% share", "2018" = "2018% share","change" ="% change") %>% gather("Year", "Percent", "2019":"2018")

bdTools$Percent <- as.numeric(unlist(str_remove_all(bdTools$Percent, "\\%")))

#Visualization of Want and Have
ggplot(bdTools, aes(fill=Year, y=Percent, x=reorder(Platform,Percent))) + 
  scale_fill_brewer(palette="Accent") +
  geom_col(position="dodge") +
  labs(y = "Percent of Respondants", x =" Big Data Tools", title = "Big Data Tools Used in DS") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip()
```

#Programming Languages in Demand for Data Scientists

```{r}
progLang <- kdTableFour[[6]]
kable(progLang) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Clean Up the table for analysis
progLang <- rename(progLang, "2019" = "2019% share", "2018" = "2018% share","change" ="% change") %>% gather("Year", "Percent", "2019":"change")

progLang$Percent <- as.numeric(unlist(str_remove_all(progLang$Percent, "\\%")))

progLang <- filter(progLang, Year == "change") %>% drop_na()

#Visualization of Want and Have
ggplot(progLang, aes(y=Percent, x=reorder(Platform,Percent))) +
  geom_col(fill = "pink") +
  labs(y = "Percent Change", x ="Programming Platforms", title = "Programming Platform Trends (2018 - 2019)") +
   theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() 
```